---
title: "EDA Diamond Prices Part 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Diamonds

Load the Diamonds data set

- How many observations are in the data set? **53,940**
- How many variables? **10**
- How many ordered factors? **3**
- What letter represents the best color? **D**

```{r, message=FALSE}
library(ggplot2)
```

```{r}
str(diamonds)
```

### 2. Price Histogram

Create a histogram of the price of all the diamonds in the diamond data set

```{r}
qplot(x = price, data = diamonds, binwidth = 100, color = I("black"), fill = I("orange"))
```

### 3. Price Histogram Summary

Describe the shape and centre of the distribution.Include summary statistics.

**The histogram represents a right skewed distribution, with a median of $2,401 and a mean of $3,933**

```{r}
summary(diamonds$price)
```

### 4. Diamond Counts

- How many diamonds cost under $500? **1,729**
- How many diamonds cost under $250? **0**
- How many diamonds cost $15,000 or more? **1,656**

```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```

### 5. Cheaper Diamonds

Explore the largest peak in the price histogram you created earlier. Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.

```{r}
qplot(x = price, data = diamonds, binwidth = 250, color = I("black"), fill = I("gold")) +
  scale_x_continuous(limits = c(0, 18823), breaks = seq(0, 18823, 1000))
```

### 6. Price by Cut Histograms

Break out the histogram of diamond prices by cut.You should have five histograms in separate panels on your resulting plot.

```{r}
qplot(x = price, data = diamonds, binwidth = 250, color = I("black"), fill = I("yellow")) +
  facet_wrap(~cut)
```

### 7. Price by Cut

- Which cut has the highest priced diamond? **Premium**
- Which cut has the lowest priced diamond? **Premium and Ideal**
- Which cut has the lowest median price? **Ideal**

```{r, message=FALSE}
library(dplyr)
```

```{r}
filter(diamonds, price == max(diamonds$price))
filter(diamonds, price == min(diamonds$price))
table(diamonds$cut)
  by(diamonds$price, diamonds$cut, summary)
```

### 8. Scales and Multiple Histograms

Add a parameter to facet_wrap so that the y-axis in the histograms is not fixed. You want the y-axis to be different for each histogram.

```{r}
qplot(x = price, data = diamonds, binwidth = 250, color = I("black"), fill = I("green")) +
  facet_wrap(~cut, scales = "free_y")
```

### 9. Price per Carat by Cut

Create a histogram of price per carat and facet it by cut. Adjust the bin width and transform the scale of the x-axis using log10.

```{r}
qplot(x = log10(price/carat + 1), data = diamonds, binwidth = 0.05, color = I("black"), fill = I("blue")) +
  facet_wrap(~cut, scales = "free_y")
```

### 10. Price Box Plots

Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

```{r}
qplot(x = color, y = price, data = diamonds, geom = "boxplot")

table(diamonds$color) 
  by(diamonds$price, diamonds$color, summary)
```

### 11. Interquartile Range

- What is the price range for the middle 50% of diamonds with color D? **911 - 4,214**
- What is the price range for the middle 50% of diamonds with color J? **1,861 - 7,695**
- What is the IQR for diamonds with the best color? **3,303**
- What is the IQR for diamonds with the worst color? **5,835**

**All figures are available in the summary tables by diamond color above**

Can also find IQR as follows:
```{r}
IQR(diamonds$price[diamonds$color == "D"])
IQR(diamonds$price[diamonds$color == "J"])
```

### 12. Price per Carat Box Plots by Color

Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.

```{r}
qplot(x = color, y = price/carat, data = diamonds, geom = "boxplot")
```

### 13. Carat Frequency Polygon

Invesitgate the weight of diamonds using a frequency polygon. 

Which carat size has a count greater than 2000? **0.3, 1.01**

```{r}
qplot(x = carat, data = subset(diamonds, !is.na(carat)), geom = "freqpoly", binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0, 5, 0.1))

filter(count(diamonds, carat, carat > 2000), n > 2000)
```

### 14. Data Wrngling in R

tidyr -a package that reshapes the layout of your data

dplyr - a package that helps you transform tidy, tabular data

### 15. Gapminder Data

Download a data set of your choice and create 2-5 plots that make use of the techniques from Lesson 3. You might use a simple histogram, a boxplot split over a categorical variable, or a frequency polygon. The choice is yours! You can find a link to the Gapminder website in the Instructor Notes.

Once you've completed your investigation, create a post in the discussions that includes:

- any questions you answered, your observations, and summary statistics
- snippets of code that created the plots
- links to the images of your plots

**Each year, countries are scored on how corrupt their public sectors are seen to be.**

**Load the Corruption Perception Index into a data frame and examine it's contents:**

```{r}
cpi <- read.table("cpi.csv" , header = TRUE, sep = ",", quote = "\"", na.strings = "")
head(cpi)
```

**Plotting CPI index of 2008 vs 2009 shows a high linear correlation of scores. This indicates most country's CPI scores did not change much from one year to the next.**

```{r}
qplot(x = X2008, y = X2009, data = subset(cpi, !is.na(X2009)), 
      color = I("black"), fill = I("red"))
```

**Comparing frequency polygons, it appears the median CPI score may be rising**

```{r, message=FALSE}
library(gridExtra)
```
```{r}
plot1 <- qplot(x = X2008, data = subset(cpi, !is.na(X2008)), geom = "freqpoly", binwidth = .5) +
  scale_x_continuous(breaks = seq(0, 10, 1))
plot2 <- qplot(x = X2009, data = subset(cpi, !is.na(X2009)), geom = "freqpoly", binwidth = .5) +
  scale_x_continuous(breaks = seq(0, 10, 1))
grid.arrange(plot1, plot2, ncol = 1)
```

**However, when comparing box plots, it appears the IQR actually decreased from 2008 to 2009**

```{r}
plot3 <- qplot(1, X2008, data = subset(cpi, !is.na(X2008)), geom = "boxplot", 
    xlab = "2008", ylab = "CPI Score") +
  scale_x_continuous(breaks = seq(0, 1, 1))
plot4 <- qplot(1, X2009, data = subset(cpi, !is.na(X2009)), geom = "boxplot", 
    xlab = "2009", ylab = "CPI Score") +
  scale_x_continuous(breaks = seq(0, 1, 1))
grid.arrange(plot3, plot4, nrow = 1)
```

**When examining a summary of each year and the IQR's, we see the median value is the same, the mean has decreased by 0.024, the IQR value is the same but has shifted lower, and the maximum score has dropped by 0.1.**

```{r}
summary(cpi$X2008)
summary(cpi$X2009)
IQR(cpi$X2008, na.rm = TRUE)
IQR(cpi$X2009, na.rm = TRUE)
```

### 16. Exploring Your Friend's Birthdays

Investigate the distribution of your friends' birth months and days.

- How many people share your birthday?
- Which month contains the most number of birthdays?
- How many birthdays are in each month?
- Which day of the year has the most number of birthdays?

**Load the facebook friends birthday data into a data frame and examine it's contents:**

```{r}
FBfriends <- read.csv("FBfriends.csv" , header = TRUE)
head(FBfriends)
```

**Separate day and month**

```{r}
FBfriends$month <- as.numeric(substr(FBfriends$Start.date, 1, 2))
FBfriends$day <- as.numeric(substr(FBfriends$Start.date, 4, 5))
```

How many people share your birthday? **none**

```{r}
sum(FBfriends$Start.date == 11/13/2016)
```

Which month contains the most number of birthdays? **June**

```{r}
qplot(x = month, data = FBfriends, binwidth = 1, color = I("black"), fill = I("orange")) +
  scale_x_continuous(breaks = seq(0, 12, 1))
```

How many birthdays are in each month?

```{r}
table(FBfriends$month)
```

Which day of the year has the most number of birthdays? **18, 27, 30**
```{r}
qplot(x = day, data = FBfriends, binwidth = 1, color = I("black"), fill = I("orange")) +
  scale_x_continuous(breaks = seq(0, 31, 1))

table(FBfriends$day)
```